---
title: "aggregate visualization data"
output: html_document
---



Package loading...
```{r}
library(ggplot2)
library(dplyr)
library(countrycode)
```



```{r, echo=FALSE}
setwd(paste0(getwd(), "/GitHub/udacity-dand-viz-project"))
getwd()
# suppress scientific notation just a hack i guess
options(scipen=999)
#list.files()
```



Let's add country code to the mobile data so that we can merge the GNI data by a common value later. Also, we'll keep only the relevant features.
```{r}
mobile <- read.csv("mobile-parameters-q2-2015-updated.csv",
                   header = TRUE, sep = ",")

mobile$Code <- countrycode(mobile$Country, "country.name", "wb")

mobile <- mobile %>% select(Country, Code, Usage.allowance, Allowance.units,
                            Validity..days., Cost, Tarrif.Currecy)

# at the end of this script before writing to csv,
# we found out that Togo's currency code is wrong messing up a few values...
mobile$Tarrif.Currecy <- as.character(mobile$Tarrif.Currecy)
index <- mobile$Country == "Togo"
mobile$Tarrif.Currecy[index] <- "XOF"
```



Now we can add the relevant exchange rate and GNI data...
```{r}
gni_usd <- read.csv("world-gni-historic.csv", header = TRUE,
                           sep = ",")
xrates_usd <- read.csv("xrates-2015.csv", header = TRUE, sep = ",")

gni_usd <- gni_usd %>% select(Country.Code, X2014)
```



Let's check if there are any errors in the currency coding in the mobile data set...
```{r}
check <- mobile %>% distinct(Tarrif.Currecy) %>% group_by(Country) %>%
  summarise(count = n()) %>% filter(count > 1)

check <- as.vector(check$Country)
check

redos <- mobile %>% filter(Country %in% check) %>%
  distinct(Tarrif.Currecy) %>% select(Tarrif.Currecy)

redos <- as.vector(unlist(redos))
redos
```



We can recode the currency coding errors to keep that data...
```{r}
index <- mobile$Tarrif.Currecy == "AFS"
mobile$Tarrif.Currecy[index] <- "AFN"

index <- mobile$Tarrif.Currecy == "BYR "
mobile$Tarrif.Currecy[index] <- "BYR"

index <- mobile$Tarrif.Currecy == "CKZ"
mobile$Tarrif.Currecy[index] <- "CZK"

index <- mobile$Tarrif.Currecy == "MGA "
mobile$Tarrif.Currecy[index] <- "MGA"

index <- mobile$Tarrif.Currecy == "GBP "
mobile$Tarrif.Currecy[index] <- "GBP"
```



Now we can merge the exchange rate and GNI data to the mobile data...
```{r}
combine <- merge(mobile, xrates_usd, by.x = "Tarrif.Currecy",
                  by.y = "Currency_Code", all.x = TRUE)

combine <- merge(combine, gni_usd, by.x = "Code",
                  by.y = "Country.Code", all.x = TRUE)
```



The names are not precise and kind of inconvenient...
```{r}
names(combine)

names(combine) <- c("country.code", "currency.code", "country.name",
                    "data.amount", "data.unit", "expiry", "cost.local", "usdxrate",
                    "gni")
```



Get rid of rows without data...
```{r}
# change empty string to NA
combine[combine == ""] <- NA

combine <- na.omit(combine)
```



Let's make sure that our features are the correct data type
```{r}
str(combine)

# as numeric from factor would assign a number corresponding to the level
# of the factor. we need as character before numeric to get the value.
combine$expiry <- as.numeric(as.character(combine$expiry))
combine$cost.local <- as.numeric(as.character(combine$cost.local))

# lost 1 data point to coercion
combine <- na.omit(combine)

str(combine)
```



Create a USD cost for each mobile plan and a monthly GNI.
```{r}
combine <- combine %>% mutate(cost.usd = cost.local/usdxrate) %>%
  mutate(gni.month = gni/12)
```



Let's convert MB to GB for 1, 7 and 30 day plans for easier aggregation. Let's check plan expiry types and counts first...
```{r}
# counts of different durations of plans
plan.counts <- combine %>% group_by(expiry) %>% summarise(count = n())

# 1, 7, AND 30 DAY PLANS HAVE THE MOST DATA.
# WE COULD FACTOR SOME OF THE REMAINING DATA (2,3,14,28) TO ADD MORE
# INSTANCES FOR 1 AND 7 DAYS BUT LET'S SEE HOW IT GOES FIRST



# filter for 1, 7, and 30 day plans (note single OR divider)
combine <- combine %>% filter(expiry == 1 | expiry == 7 | expiry == 30)

# logical vector of rows with "MB" as data unit
#mb <- combine$data.unit == "MB" & (combine$expiry == 1 | combine$expiry == 7 | combine$expiry == 30)
mb <- combine$data.unit == "MB"

# CHECK SELECTION BEFORE
#combine[mb, ]

# slice df by "GB" and "1" rows and assign new units and amounts
combine[mb, ]$data.amount <- (combine[mb, ]$data.amount / 1000)
combine[mb, ]$data.unit <- "GB"

# CHECK SELECTION AFTER
#combine[mb, ]
```



Now let's group by country and plan/expiry type to get total data amount and total cost for each plan duration. From here we can get an average cost in USD per GB for each plan duration for a given country.
```{r}
# need to groupby 'gni.month' and 'country.code' here or it gets removed in the chain
final_df <- combine %>%
  group_by(country.name, expiry, country.code, gni.month, currency.code) %>%
  summarise(total.amount = sum(data.amount), total.cost = sum(cost.usd)) %>%
  mutate(cost.usd.per.gb = total.cost/total.amount)
```



Some kind of benchmark usage cost would be good to establish a typical monthly bill across plans/expiry.  Some sources that say the average person consumes about 2GB of mobile data (albeit in the US but this also doesn't include mobile wifi which is apparently close to 8GB per month) so a cost for this average consumption per plan/expiry would be nice.  We could also create another feature representing the proportion of the 2GB monthly consumption cost as a fraction of monthly GNI per person in that country.  This will really help to establish a relative measure for the choropleth map.
```{r}
final_df <- final_df %>%
  mutate(cost.usd.bench.2gb.month = 2*cost.usd.per.gb) %>%
  mutate(percent.bench.income = 100*cost.usd.bench.2gb.month/gni.month) %>%
  select(-total.amount, -total.cost)
```



Let's make sure our relative measure domain is reasonable for a choropleth map...
```{r}
summary(final_df)
ggplot(aes(x = percent.bench.income), data = final_df) +
  geom_histogram()
# we have one major outlier, let's see all high values...
check <- subset(final_df, percent.bench.income > 50)
```



Looking at some of the costs in the original mobile data it seems there may have been some bad data entry so I'm going to exclude anything above %50...
```{r}
final_df <- subset(final_df, percent.bench.income <= 50)
summary(final_df)
ggplot(aes(x = percent.bench.income), data = final_df) +
  geom_histogram()
```



Write data to csv for d3 visualization...
```{r}
ordering <- c("country.name", "country.code", "expiry", "cost.usd.per.gb",
              "cost.usd.bench.2gb.month", "percent.bench.income", "gni.month")

final_df <- final_df[, ordering]

write.csv(final_df, file = "mobile-world-data.csv", row.names = FALSE)
```
