---
title: "aggregate visualization data"
output: html_document
---



Package loading...
```{r packages}
library(ggplot2)
library(dplyr)
library(tidyr)
library(countrycode)
```


Setting local directory and suppressing scientific notation...
```{r set_directory, echo=FALSE}
# make sure local path is same as this file
getwd()
#setwd(paste0(getwd(), "/GitHub/udacity-dand-viz-project/data"))
list.files()
# suppress scientific notation, just a hack i guess...
options(scipen=999)
```


Let's import the mobile data set and add country code so that we can merge the GNI data by a common value later.
```{r mobile_import}
mobile <- read.csv("mobile-parameters-q4-2014.csv",
                   header = TRUE, sep = ",", strip.white = TRUE)

mobile$Code <- countrycode(mobile$Country, "country.name", "wb")
```


Let's add the relevant GNI data...
```{r gni_xrates_import}
gni_usd <- read.csv("world-gni-historic.csv", header = TRUE,
                           sep = ",")
```


Any missing 2014 GNI data is replaced by either 2013 or in a few rare cases 2012...
```{r gni_fill_missing}
nas <- is.na(gni_usd$X2014)
gni_usd[nas, ]$X2014 <- gni_usd[nas, ]$X2013

nas <- is.na(gni_usd$X2014)
gni_usd[nas, ]$X2014 <- gni_usd[nas, ]$X2012

# still missing a good chunk of data
nas <- is.na(gni_usd$X2014)
sum(nas)
# for these countries
as.vector(unlist(gni_usd[nas, ]$Country.Code))

gni_usd <- gni_usd %>% select(Country.Code, X2014)
```


Now we can merge the exchange rate and GNI data to the mobile data...
```{r combine_data_sets}
combine <- merge(mobile, gni_usd, by.x = "Code",
                  by.y = "Country.Code", all.x = TRUE)
```


Let's get rid of some columns...
```{r prune_columns}
combine <- combine %>% select(-Region, -ISP, -Pre.or.Post.paid, -Usage.Allowance..MB.,
                              -Tariff.currency, -Monthly.cost..specified.currency.,
                             -Pack.Cost..specified.currency., -Voice.mins.included.,
                             -Validity..days., -Monthly.cost..US.., -Pack.Cost..US..,
                             -Standard.Xrate, -VAT.Rate, -Date.tariff.recorded,
                             -Source.URL)
```


Rename and format columns...
```{r rename_and_format_columns}
names(combine)
names(combine) <- c("country.code", "country.name", "usage.amount",
                    "usage.unit", "cost.usd.gb", "gni")

str(combine)

#combine$region <- as.character(combine$region)
combine$country.name <- as.character(combine$country.name)

combine$usage.amount <- as.character(combine$usage.amount)
combine$usage.unit <- as.character(combine$usage.unit)

#combine$expiry <- as.numeric(as.character(combine$expiry))
combine$cost.usd.gb <- as.numeric(as.character(combine$cost.usd.gb))
combine$gni <- as.numeric(as.character(combine$gni))
```


How many unlimited plans?
```{r unlimited_count}
unlimited <- combine %>% filter(usage.amount == "Unlimited")

# add a count for each country in the unlimited df,
# we can add counts for each country to the combine df LATER
unlimited <- unlimited %>% group_by(country.name) %>% summarise(count = n())

# remove unlimited plans from the combine df
combine <- combine %>% filter(usage.amount != "Unlimited")
```


How many have bad usage units...
```{r unit_count}
bad.unit <- combine %>% filter(usage.unit != "MB" & usage.unit != "GB")

# it looks like some countries have hourly plans with Taiwan having a 'days' plan
# the Taiwan data looks bad, let's exlude, also let's make another df for hourly plans
hourly <- combine %>% filter(usage.unit %in% c("hour", "hours"))
hourly <- hourly %>% group_by(country.name) %>% summarise(count = n())

# remove hourly plans from the combine df
combine <- combine %>% filter(!(usage.unit %in% c("hour", "hours", "days")))
```


Double check our amounts and units...
```{r double_check_amount_and_unit_types}
table(combine$usage.amount)
table(combine$usage.unit)
```


Now that we have country name together with GNI data we can add a few GNIs manually from the web...
```{r combine_manual_gni}
nas <- is.na(combine$gni)
# missing data
missing <- combine[nas, ]
# and the counts
counts <- missing %>% group_by(country.name) %>% summarise(n())

index <- combine$country.code == "AGO"
combine$gni[index] <- 6670

index <- combine$country.code == "ROM"
combine$gni[index] <- 18060

index <- combine$country.code == "SOM"
combine$gni[index] <- 125

index <- combine$country.code == "SYR"
combine$gni[index] <- 1784

index <- combine$country.code == "TWN"
combine$gni[index] <- 22598

index <- combine$country.code == "ZAR"
combine$gni[index] <- 440

# no gni data for these countries
nas <- is.na(combine$gni)
# no more missing data
sum(nas)
```


Let's create a monthly gni column...
```{r monthly_gni}
combine$gni.month <- combine$gni / 12
```


See if we have any missing cost per gb data...
```{r missing_cost}
sum(is.na(combine$cost.usd.gb))
```


Now let's group by country and usage unit to get the average cost in USD for GB or MB based plans...
```{r final_df_costpergb}
# need to groupby 'gni.month' and 'country.name' here or it gets removed in the chain
final_df <- combine %>% group_by(country.code, usage.unit,
                                 country.name, gni.month) %>%
  summarise(count = n(), total.cost.gb.usd = sum(cost.usd.gb)) %>%
  mutate(cost.usd.per.gb = total.cost.gb.usd/count) %>%
  select(-total.cost.gb.usd, -count)
```


Let's spread this data into one row per country and name accordingly...
```{r spread_columns}
final_df <- spread(final_df, usage.unit, cost.usd.per.gb)

names(final_df)
names(final_df) <- c("country.code", "country.name", "gni.month",
                    "gb.cost.gb", "gb.cost.mb")
```


We have average cost in USD per GB but this doesn't quite paint an accurate picture mobile data accessibility.  For this we can compare the typical monthly usage (2GB) bill divided by the montly GNI per capita.  The 2GB approximate average monthly usage comes from a few online sources.  In the US, the average person consumes about 2GB of mobile data (although this doesn't include mobile wifi as in free hotspots at starbucks which is apparently close to an additional 8GB per month). This relative measure of mobile data accessibility will really help in bringin the choropleth map to life...
```{r final_df_percent_income}
final_df <- final_df %>%
  mutate(percent.income.gb = 100*2*gb.cost.gb/gni.month) %>%
  mutate(percent.income.mb = 100*2*gb.cost.mb/gni.month)
```


Let's make sure our domains are reasonable for a choropleth map...
```{r final_df_domain_check}
ggplot(aes(x = gb.cost.gb), data = final_df) +
  geom_histogram()

ggplot(aes(x = gb.cost.mb), data = final_df) +
  geom_histogram()

# we have a major outlier in the mb costs...
check <- subset(final_df, gb.cost.mb > 100)
# based upon this finding it looks like the original csv had some bad MB to GB
# encoding for one row in Syria which was changed manually
```


Let's cut cost per GB and cost as percent of income into buckets for better visualization in d3...
```{r final_df_buckets}
summary(final_df)

final_df$gb.cost.gb.bucket <- cut(final_df$gb.cost.gb,
                                      c(0, 3, 6, 10, 15, 25, +Inf),
                                      labels = c('0-3', '3-6', '6-10', '10-15',
                                                 '15-25', '25-'))
final_df$gb.cost.gb.bucket <- as.character(final_df$gb.cost.gb.bucket)

final_df$gb.cost.mb.bucket <- cut(final_df$gb.cost.mb,
                                      c(0, 7, 15, 30, 50, 80, +Inf),
                                      labels = c('0-7', '7-15', '15-30', '30-50',
                                                 '50-80', '80-'))
final_df$gb.cost.mb.bucket <- as.character(final_df$gb.cost.mb.bucket)



final_df$percent.income.gb.bucket <- cut(final_df$percent.income.gb,
                                      c(0, 1, 3, 10, 30, 50, +Inf),
                                     labels = c('0-1', '1-3', '3-10',
                                                '10-30', '30-50', '50-'))
final_df$percent.income.gb.bucket <- as.character(final_df$percent.income.gb.bucket)

final_df$percent.income.mb.bucket <- cut(final_df$percent.income.mb,
                                      c(0, 3, 7, 30, 50, 80, +Inf),
                                     labels = c('0-3', '3-7', '7-30', '30-50',
                                                 '50-80', '80-'))
final_df$percent.income.mb.bucket <- as.character(final_df$percent.income.mb.bucket)
```


Add unlimited and hourly plan counts?
```{r add_unlimited_hourly}
#
```


While debugging in javascript, a few country codes don't match up.  A few need to be changed to match up with the geojson file...
```{r}
# Hong Kong HKG and Singapore SGP don't have indepenent map areas in the geojson file...
index <- final_df$country.code == "ZAR"
final_df$country.code[index] <- "COD"
index <- final_df$country.code == "ROM"
final_df$country.code[index] <- "ROU"
```


Write data to csv for d3 visualization...
```{r final_df_write_csv}
names(final_df)

# round numbers
final_df$gni.month <- round(final_df$gni.month, 2)
final_df$gb.cost.gb <- round(final_df$gb.cost.gb, 2)
final_df$gb.cost.mb <- round(final_df$gb.cost.mb, 2)
final_df$percent.income.gb <- round(final_df$percent.income.gb, 2)
final_df$percent.income.mb <- round(final_df$percent.income.mb, 2)

write.csv(final_df, file = "mobile-world-data.csv", row.names = FALSE)
```
