---
title: "aggregate visualization data"
output: html_document
---



Package loading...
```{r}
library(ggplot2)
library(dplyr)
library(countrycode)
```



```{r, echo=FALSE}
setwd(paste0(getwd(), "/GitHub"))
getwd()
# suppress scientific notation just a hack i guess
options(scipen=999)
#list.files()
```



Let's add country code to the mobile data so that we can merge the GNI data by a common value later. Also, we'll keep only the relevant features.
```{r}
mobile <- read.csv("mobile-parameters-q2-2015-updated.csv", header = TRUE, sep = ",")
mobile$Code <- countrycode(mobile$Country, "country.name", "wb")
mobile <- mobile %>% select(Country, Code, Usage.allowance, Allowance.units,
                            Validity..days., Cost, Tarrif.Currecy)
```



Now we can add the relevant exchange rate and GNI data...
```{r}
gni_usd <- read.csv("world-gni-historic.csv", header = TRUE,
                           sep = ",")
xrates_usd <- read.csv("xrates-2015.csv", header = TRUE, sep = ",")

gni_usd <- gni_usd %>% select(Country.Code, X2014)
```



Let's check if there are any errors in the currency coding in the mobile data set...
```{r}
check <- mobile %>% distinct(Tarrif.Currecy) %>% group_by(Country) %>%
  summarise(count = n()) %>% filter(count > 1)

check <- as.vector(check$Country)
check

redos <- mobile %>% filter(Country %in% check) %>%
  distinct(Tarrif.Currecy) %>% select(Tarrif.Currecy)

redos <- as.vector(unlist(redos))
redos
```



We can recode the currency coding errors to keep that data...
```{r}
index <- mobile$Tarrif.Currecy == "AFS"
mobile$Tarrif.Currecy[index] <- "AFN"

index <- mobile$Tarrif.Currecy == "BYR "
mobile$Tarrif.Currecy[index] <- "BYR"

index <- mobile$Tarrif.Currecy == "CKZ"
mobile$Tarrif.Currecy[index] <- "CZK"

index <- mobile$Tarrif.Currecy == "MGA "
mobile$Tarrif.Currecy[index] <- "MGA"

index <- mobile$Tarrif.Currecy == "GBP "
mobile$Tarrif.Currecy[index] <- "GBP"
```



Now we can merge the exchange rate and GNI data to the mobile data...
```{r}
combine <- merge(mobile, xrates_usd, by.x = "Tarrif.Currecy",
                  by.y = "Currency_Code", all.x = TRUE)

combine <- merge(combine, gni_usd, by.x = "Code",
                  by.y = "Country.Code", all.x = TRUE)
```



The names are not precise and kind of inconvenient...
```{r}
names(combine)

names(combine) <- c("country.code", "currency.code", "country.name",
                    "data.amount", "data.unit", "expiry", "cost.local", "usdxrate",
                    "gni")
```



Get rid of rows without data...
```{r}
# change empty string to NA
combine[combine == ""] <- NA

combine <- na.omit(combine)
```



Let's make sure that our features are the correct data type
```{r}
str(combine)

# as numeric from factor would assign a number corresponding to the level
# of the factor. we need as character before numeric to get the value.
combine$expiry <- as.numeric(as.character(combine$expiry))
combine$cost.local <- as.numeric(as.character(combine$cost.local))

# lost 1 data point to coercion
combine <- na.omit(combine)

str(combine)
```



Create a USD cost for each mobile plan, a yearly cost, and a yearly percent of GNI cost...
```{r}
combine <- combine %>% mutate(cost.usd = cost.local/usdxrate)#, 
                              #cost.year.usd = cost.usd*(365/expiry),
                              #cost.year.perc.gni = cost.year.usd/gni)
# THE BY YEAR COSTS ARE TOO PREMATURE HERE PUT THEM BELOW!!!!!!!

```



Let's convert MB to GB for 1, 7 and 30 day plans for easier aggregation. Let's check plan expiry types and counts first...
```{r}
# counts of different durations of plans
plan.counts <- combine %>% group_by(expiry) %>% summarise(count = n())

# 1, 7, AND 30 DAY PLANS HAVE THE MOST DATA.
# WE COULD FACTOR SOME OF THE REMAINING DATA (2,3,14,28) TO ADD MORE
# INSTANCES FOR 1 AND 7 DAYS BUT LET'S SEE HOW IT GOES FIRST



# filter for 1, 7, and 30 day plans (note single OR divider)
combine <- combine %>% filter(expiry == 1 | expiry == 7 | expiry == 30)

# logical vector of rows with "MB" as data unit
#mb <- combine$data.unit == "MB" & (combine$expiry == 1 | combine$expiry == 7 | combine$expiry == 30)
mb <- combine$data.unit == "MB"

# CHECK SELECTION BEFORE
#combine[mb, ]

# slice df by "GB" and "1" rows and assign new units and amounts
combine[mb, ]$data.amount <- (combine[mb, ]$data.amount / 1000)
combine[mb, ]$data.unit <- "GB"

# CHECK SELECTION AFTER
#combine[mb, ]
```



Now let's group by country and plan/expiry type and get total data amount and total cost...
```{r}
# need to groupby 'gni' here or it gets removed in the chain
test <- combine %>% group_by(country.name, expiry, gni) %>%
  summarise(total.amount = sum(data.amount), total.cost = sum(cost.usd))

test <- test %>% mutate(cost.usd.per.gb = total.cost/total.amount)

test <- test %>% mutate(cost.year.usd = cost.usd.per.gb*(365/expiry),
                        cost.year.perc.gni = cost.year.usd/gni)


```




Write data to csv for d3 visualization...
```{r}
write.csv(combine, file = "mobile-world-data.csv", row.names = FALSE)
```
